
\section{A connection to the Kneser graphs $\KG(2n-3,n-2)$}

In this section we want to give an alternative proof of the MDS main conjecture for $n\leq p$ using essentially the same means as before and some known facts about the eigenvalues of Kneser graphs.

We recall the interpolation lemma for the tangent polynomials which states that for disjoint subsets $A,B\setleq \cA$ ($\card{A}=t+2$, $\card{B}=n-2$ and w.l.o.g.~$t+n < \card{\cA}$ using the dualisation argument)
we have that

$$
\sum_{a\in A}{P(\set a\setjoin B)\prod_{z\in A\setminus \set a}{\det(z,\set a\setjoin B)}^{-1}}=0\label{mds-tan-poly-interpol}
$$

(this is \autoref{mds-abc-lemma} for $r=1$). Now, we define the interpolation terms as

$$
I(C,D):=P(C)\prod_{d\in D}{{\det(d,C)}^{-1}}
$$
  
where $C,D\setleq \cA$, $\card{C}=n-1$, $\card{D}=t+1$.
If we now rewrite the equation~\theref{mds-tan-poly-interpol} % fix this : use equation environment -> better definition if \label, \autoref, \theref
one sees that we get a linear system of
$$
\binom{\card\cA}{t+2,n-2,\card\cA-n-t}
$$
equations in
$$
\binom{\card\cA}{t+1,n-1,\card\cA-n-t}
$$
variables, namely
$$% bug : use \text
\sum_{a\in A}{I(A\setminus\{a\}, B\setjoin\{a\})}=0\text{ for $A\in\binom{\cA}{t+2}$, $B\in\binom{\cA}{n-2}$, $A\setmeet B=\emptyset$}.\label{mds-interpol-sys}
$$
Moreover, this system decomposes into $\binom{\card\cA}{n+t}$ independent components on which $A\setjoin B=E$ where $E\in\binom{\cA}{t+n}$ is constant.

Note that for the critical case $t=n-3$ there are the same number of variables as equations.
We will now show that in this case, each such component can be interpreted as the adjacency matrix of the Kneser graph $\KG(2n-3,n-2)$ and deduce that it is regular when interpreted as a matrix in $\ints/p\ints$ where $p\leq n$ using a fact about the eigenvalues of Kneser graphs.

\begin{lemma}
    Assume $t=n-3$ (i.e. $\card{\cA}=q+2$).
    Let $E\in\binom{\cA}{t+n}$. Then the component of the linear system~\theref{mds-interpol-sys}
    on which $A\setjoin B=E$ is the adjacency matrix of the graph $\KG(2n-3,n-2)$.
\end{lemma} 

\begin{proof}
    Let $x_D:=I(D,E\setminus D)$. Then for each $(n-2)$-set $C\setleq E$ there is an equation
    $$
    \sum_{D:C\setmeet D=\emptyset}{x_D}=0
    $$
    and this is clearly the system $Ax=0$ where $A=(a_{CD})$ is the adjacency matrix of $\KG(2n-3,n-2)$ with
    $$
    a_{CD}=
    \begin{cases}
        1 &: C\setmeet D=\emptyset\\
        0 &: \otherwise
    \end{cases}.
    $$
    and $x=(x_D)$.
\end{proof}

For $n\leq p$ one derives easily a contradiction from this using a well-known fact about the eigenvalues of Kneser graphs.

\begin{lemma}[graph spectrum of Kneser graphs]
  Let $2k\leq n$. Then the eigenvalues in $\rats$ of the adjacency matrix of the Kneser graph $\KG(n,k)$ are
  $$
    \lambda_j:={(-1)}^j\binom{n-k-j}{k-j}
  $$
  for $j=0,\ldots,k$ where $\lambda_j$ has geometric and algebraic multiplicity $\binom{n}{j}-\binom{n}{j-1}$.
\end{lemma}

\begin{proof}
  Set 
  $$
  V_j:=\dirsum_{\substack{A\setleq n\\ \card{A}=j}}{\rats A}
  $$
  for $j=0,\ldots,n$.
  Moreover, define linear mappings $\phi^l_{ij}:V_j\to V_i$ by
  $$
  \phi^l_{ij}(B) = \sum_{\substack{A\setleq n\\ \card{A\setmeet B}=l, \card{A}=i}}{A}
  $$
  and linear continuation.
  For these we may compute that
  \begin{align*}
      \phi^l_{ij}\compose\phi^m_{jk}(C)
      &= \sum_{\substack{B\setleq n\\ \card{B\setmeet C}=m, \card{B}=j}}{\sum_{\substack{A\setleq n\\ \card{A\setmeet B}=l, \card{A}=i}}{A}}\\
      & = \sum_{\substack{A\setleq n\\ \card{A}=i}}{\card{\{B\setleq n: \card{B}=j\lgand\card{A\setmeet B}=l\lgand\card{B\setmeet C}=m\}}{A}}\\
      & = \sum_{\nu\in\nats}{\sum_{\substack{A\setleq n\\ \card{A}=i, \card{A\setmeet C}=\nu}}{\left(\sum_{\mu=0}^{\nu}{\binom{i-\nu}{l-\mu}\binom{k-\nu}{m-\mu}\binom{\nu}{\mu}\binom{n-i-k+\nu}{j-l-m+\mu}}\right)A}}\\
      & = \sum_{\nu\in\nats}{\left(\sum_{\mu=0}^{\nu}{\binom{i-\nu}{l-\mu}\binom{k-\nu}{m-\mu}\binom{\nu}{\mu}\binom{n-i-k+\nu}{j-l-m+\mu}}\right)\phi^{\nu}_{ik}(C)}
  \end{align*}
  where we distinguish the sets $A$ by the size $\nu$ of $\card{A\setmeet C}$ and then count the sets $B$ which intersect with $A$ and $C$ in the right amount (here $\mu$ is $\card{A\setmeet B\setmeet C}$).
  This implies the identities
  $$
  \phi^j_{ij}\compose\phi^k_{jk} = \binom{i-k}{j-k}\phi^k_{ik},
  $$
  $$
  \phi^i_{ij}\compose\phi^j_{jk} = \binom{k-i}{j-i}\phi^i_{ik},
  $$
  and
  $$
  \phi^0_{ij}\compose\phi^k_{jk} = \binom{n-i-k}{j-k}\phi^0_{ik}.\label{kglemma:eq1}
  $$
  From the first identity we deduce that $\im(\phi^k_{ik})\setleq\im(\phi^j_{ij})$ for $k\leq j\leq i$. Similarly, from the third we get $\im\phi_{ik}^0\setleq\im\phi_{ij}^0$ for $k\leq j\leq n-i$.
  Moreover, we need the identities
  $$
  \phi^j_{ij}=\sum_{k=0}^{j}{{(-1)}^k\phi^0_{ik}\compose\phi^k_{kj}}.\label{kglemma:eq2}
  $$
  and
  $$
  \phi_{ij}^0=\sum_{k=0}^j{{(-1)}^k\phi_{ik}^k\compose\phi_{kj}^k}.\label{kglemma:eq3}
  $$
  The equation~\autoref{kglemma:eq3}
  follows from \autoref{kglemma:eq2}
  by composing with the linearization of the duality map $A\mapsto n\setminus A$ from the left and \autoref{kglemma:eq2}
  can be shown by
  \begin{align*}
      \left(\sum_{k=0}^{j}{{(-1)}^k\phi^0_{ik}\compose\phi^k_{kj}}\right)(C)
      & = \sum_{k=0}^{j}\sum_{\substack{A\setleq n\\ \card{A}=i}}{{(-1)}^k\card{\{B\setleq
              n:\card{B}=k\lgand  B\setleq C\setminus A\}}A}\\
      & = \sum_{\substack{A\setleq n\\ \card{A}=i}}{\delta_{\emptyset,C\setminus A}A} = \phi^j_{ij}(C)\text{.}
  \end{align*}
  Now \autoref{kglemma:eq2}
  implies $\im\phi_{ij}^j\setleq\Setjoin_{k=0}^j{\im\phi_{ik}^0}=\im\phi_{ij}^0$ ($j\leq n-i$) and \autoref{kglemma:eq3}
  implies that $\im\phi_{ij}^0\setleq\Setjoin_{k=0}^j{\im\phi_{ik}^k}=\im\phi_{ij}^j$ ($j\leq i$) such that $\im\phi_{ij}^0=\im\phi_{ij}^j$ ($j\leq i\leq n-j$).
  
  We shall now prove that $\phi^j_{ij}$ is injective if $j\leq i\leq n-j$ which can be done by applying the `dual' equation of \autoref{kglemma:eq2}
  \begin{align*}
      \id=\phi^j_{jj}
      & \stackrel{\autoref{kglemma:eq2}}{=}
      \sum_{k=0}^{j}{{(-1)}^k\phi^k_{jk}\compose\phi^0_{kj}}\\
      & \stackrel{\autoref{kglemma:eq1}}{=} \left(\sum_{k=0}^{j}{\frac{{(-1)}^k}{\binom{n-j-k}{i-j}}\phi^k_{jk}\compose\phi^0_{ki}}\right)\compose\phi^j_{ij}\text{.}
  \end{align*}
  Thus it is shown that $\phi^j_{ij}$ is an embedding when $j\leq i\leq n-i$. Let $k\leq \floor{n/2}$, then define for $j=0,\ldots,k$ the space $U_i$ as the orthogonal complement of $\im\phi_{k,j-1}^{j-1}$ in $\im\phi_{kj}^j$ (we now that $\im\phi_{k-1,j-1}^{j-1}\setleq\phi_{kj}^j$). We thus have an orthogonal decomposition
  $$
  \dirsum_{j=0}^k{U_j}=\im\phi_{k0}^0=V_k,
  $$
  which turns out to be the decomposition into eigenspaces $U_j$ of eigenvalue $\lambda_j={(-1)}^j\binom{n-k-j}{k-j}$ and dimension $\binom{n}{j}-\binom{n}{j-1}$.

  The formula for the dimension of $U_j$ follows from the fact that $\phi_{kj}^j$ are embeddings. We are left to prove the eigenspace property.

  So take $u\in U_j$ and write $u=\phi_{kj}^j(v)$. As $j\leq k\leq n-j$ it holds that $\im\phi_{ki}^i=\im\phi_{ki}^0$ for $i=0,\ldots,j-1$ and thus we have by definition of $U_j$
  $$
  0=\dual{(\phi_{ki}^0)}\compose\phi_{kj}^j(v)=\phi_{ik}^0\compose\phi_{kj}^j(v)\stackrel{\autoref{kglemma:eq1}}{=}\binom{n-i-j}{k-j}\phi_{ij}^0(v)
  $$
  for $i<j$, i.e. $\phi_{ij}^0(v)=0$ (as $k\geq j$, $n-i-k\geq 0$). But dually to the statement about the images we have $\ker\phi_{ij}^0=\ker\phi_{ij}^i$, so $\phi_{ij}^i(v)=0$.
  Lastly, we have that
  $$
  \phi_{kk}^0\compose\phi_{kj}^j\stackrel{\autoref{kglemma:eq1}}{=}\binom{n-k-j}{k-j}\phi_{kj}^0\stackrel{\autoref{kglemma:eq3}}{=}\binom{n-k-j}{k-j}\sum_{i=0}^j{{(-1)}^i\phi_{ki}^i\compose\phi_{ij}^i},
  $$
  which gives $\phi_{kk}^0(u)=\binom{n-k-j}{k-j}u$ as desired ($\phi_{kk}^0$ is the adjacency operator of $\KG(n,k)$).
\end{proof}

From this we get a direct proof of corollary \autoref{mds-bound-n-leq-p} as for $d\leq p$ there are no eigenvalues divisible by $p$ and thus the system of equations~\theref{mds-interpol-sys}
must be a regular matrix in the case $t=n-3$. Thus it would follow that $I(C,D)=0$ for all $C,D\setleq \cA$, $\card C=n-1$, $\card{D}=n-2$ which contradicts the fact that all these expressions must be units in $\field{q}$ (by definition).
%% false
It is also possible to do the above argument without the knowledge of the previous lemma. To do this, note that $S_{2n-3}$ acts on the kernel of the adjacency matrix $M$ of $\KG(2n-3,n-2)$ in $\field{q}$ (actually, it is well-known that this is the automorphism group of $\KG(2n-3,n-2)$, but we do not need that either). Assume that $d\leq p$ and that $M$ is singular. Then consider a nontrivial element of its kernel $c$. Summing the images of $c$ under all elements of $S_{2n-3}$ gives a vector which has $(n-1)!(n-2)!$ times the sum of all components of $c$ in each component. By the assumption, this cannot be the zero vector since $(n-1)!(n-2)!$ is a unit mod $p$. But then it follows that $n-1\equiv 0$ mod $p$ (plugging this vector in the equation of $M$) and thus $n>p$ which is a contradiction.   

\begin{remark}
    Actually, we do not need for a contradiction that the adjacency matrix of $\KG(2n-3,n-2)$ is regular in $\field{q}$ --- we only need that its kernel contains no vector with only non-zero components.
\end{remark}









