\message{ !name(On The Routh-Hurwitz Theorem.tex)}\documentclass[10pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{enumerate}
\input{math.tex}

%\usepackage{xparse}
%\let\rsum\sum
%\RenewDocumentCommand\sum{mo}{%sum command with optional upper index
%  \IfNoValueTF{#2}
%    {\rsum_{#1}}
%    {\rsum_{#1}^{#2}}%
%}

\begin{document}

\message{ !name(On The Routh-Hurwitz Theorem.tex) !offset(-3) }


\begin{definition}[Algebraic winding number]
Let $\gamma:[a,b]\to\complex\setminus\{0\}$ be a continuous curve ($a,b\in\reels,a\leq b$). Then there exists a unique (modulo $2\pi$) continuous function $\arg{\gamma}$ such that 
\begin{equation}
\gamma = \abs{\gamma}\exp(\i\arg{\gamma}) = \abs{\gamma}\sgn{\gamma}
\end{equation}
The \emph{algebraic winding number} of $\gamma$ with respect to the origin 0 is then defined as
\begin{equation}
\wnumb_0(\gamma) := \rest{\arg{\gamma}}^b_a = \arg{\gamma(b)}-\arg{\gamma(a)}\text{.}
\end{equation}

\end{definition}

%% for rectifiable $\sgn(\gamma)$
\begin{remark}
If $\sgn{\gamma}:I\to\complex$ is a rectifiable function, one has the identity
\begin{equation}
\wnumb_0(\gamma)=\frac{1}{2\pi\i}\int_{\sgn{\gamma}}{\frac{1}{z}}\d{z}\text{.}
\end{equation}
\end{remark}

A natural idea is now to store the rotational information (i.e. the information of the function $\arg{\gamma}$) in another function which then can be analyzed using generalized \textsc{Sturm} chains and \textsc{Cauchy} indices. However, the 'function' which is introduced may have some gaps in its domain.

\begin{definition}[\textsc{Cauchy} function] Let $\gamma:I\to\complex\setminus\{0\}$ be a continuous curve for a compact interval $I$. The \textsc{Cauchy} \emph{function} $\cfunc(\gamma)$ of $\gamma$ is then defined as
\begin{equation}
\cfunc(\gamma):I\to\reels\setminus\{x\in I:\Re{\gamma(x)}=0\},x\mapsto\frac{\Im{\gamma(x)}}{\Re{\gamma(x)}}\text{.}
\end{equation}
Any partially defined function which can be obtained in such manner is called a \textsc{Cauchy} \emph{function} over $I$.
\end{definition}


%%% unstable
Now, one needs to redefine the operations standard arithmetic operations for such partial functions.

\begin{definition}[\textsc{Cauchy} function inverse]
Let $f$ be a \textsc{Cauchy} be a function on the compact interval $I$ corresponding to the path $\gamma:I\to\complex\setminus\{0\}$. Then the \textsc{Cauchy} function $1/f$ is defined as
\begin{equation}
1/f:\dom(f)\setminus f^{-1}\{0\}\unify \intersection_{n\in\nats}f^{-1}[\reels\setminus (-n,n)]\to\reels,x\mapsto\lim_{\xi\to x}{\frac{1}{f(x)}}
\end{equation}
which gives the \textsc{Cauchy}-function corresponding to the path $\i\conj{\gamma}$.
\end{definition}

\begin{definition}[\textsc{Cauchy} function of the product path]
Let $f,g$ be \textsc{Cauchy} be a function on the compact interval $I$ corresponding to paths $\gamma,\eta:I\to\complex\setminus\{0\}$. Then the \textsc{Cauchy} function $f\star g=\frac{f+g}{1-fg}$ is defined as
...
which gives the \textsc{Cauchy}-function corresponding to the path product $\gamma\eta$.
\end{definition}

%% end unstable

%%% give a characterization of Cauchy functions

%% first definition needed

\begin{definition}[maximal rotational interval, rotational domain of a partial function]
Let $f:J\to\reels$ be continuous where $J\subset I$ is open in the subspace $I$ for a compact interval $I\subset\reels$. Let $\{I_n\}_n$ be the partition of $J$ into maximal (with respect to inclusion) intervals (which exists as $J$ lies relatively open in $I$). An interval $I_i\in\{I_n\}_n$ is called \emph{rotational} if it is not an interval of the form $(a,b)$ ($a,b\in I, a<b$) which satisfies
\begin{equation}
\lim_{\xi\downto a}{f(\xi)}=\lim_{\xi\upto b}{f(\xi)}=\pm\infty\text{.}
\end{equation}
The \emph{rotational domain} of $f$ is defined as $\conv\left(\union\{\bar{I}_i\}_i\right)$ where $\{\bar{I}_i\}_i$ are all rotational intervals among $\{I_n\}_n$.
\end{definition}

%% the characterization of Cauchy functions

\begin{lemma}[Characterization of \textsc{Cauchy} functions]
Let $I\subset\reels$ be a compact interval and $J\subset I$ be a an open set in the subspace topology $I$. Then for a continuous function $f:J\to\reels$ the following two are equivalent
\begin{enumerate}
\item The function $f$ is a \textsc{Cauchy} function over $I$.
\item Let $\{I_n\}_{n\in\nats}$ be the partition of $J$ into maximal (with respect to inclusion) intervals (which exists as $J$ lies relatively open in $I$). Then there are only finitely many rotational intervals $\{\bar{I}_i\}_{i=0}^l$ among them.
\item There exist continuous functions $g,h:I\to\reels$ with no common zero $\xi\in I$ with $g(\xi)=h(\xi)$ such that $\rest{\frac{g}{h}}_{\dom(f)}=f$ (especially the expression on the left side is well-defined on $\dom(f)$).
\end{enumerate}
\end{lemma}

\begin{proof}
\textbf{$1\implies 2$.}
Let $f$ be a \textsc{Cauchy} function corresponding to the path $\gamma:I\to\complex\setminus\{0\}$.
Now assume there would be infinitely many maximal rotational intervals. Then at most two of them are half-open intervals (at most one at each bound $a$ and $b$). Thus let $\{\bar{I}_i\}_{i\in\nats}\subset\powerset(\dom(f))=I\setminus\{x\in I:\Re{\gamma(x)}=0\}$ be all maximal open rotational intervals.
Let $\{L_i\}=\{(\hat{a}_i,\hat{b}_i)\}\subset\{\bar{I}_i\}_{i\in\nats}$ be the set of maximal open rotational intervals for which
\begin{equation}
\lim_{\xi\downto \hat{a}_i}{f(\xi)}=-\infty \vee \lim_{\xi\upto \hat{b}_i}{f(\xi)}=\infty
\end{equation}
and $\{R_i\}=\{(\check{a}_i,\check{b}_i)\}\subset\{\tilde{I}_i\}_{i\in\nats}$ such that
\begin{equation}
\lim_{\xi\downto \hat{a}_i}{f(\xi)}=\infty \vee \lim_{\xi\upto \hat{b}_i}{f(\xi)}=-\infty\text{.}
\end{equation}
(The '$\vee$' is just needed to include the half-open intervals at the boundary of $I$ which might exist.)
It is then clear that in each interval $L_i$ which is not half-open the curves $\gamma$ and $\sgn{\gamma}$ rotates the angle $\pi$ around $0$ staying in one of the half planes $\{z\in\complex:\Re(z)\geq 0\}$ or $\{z\in\complex:\Re(z)\leq 0\}$, counterclockwise. In the intervals $R_i$ which is not half-open happens the same but the angle of rotation is $-\pi$ (clockwise rotation). Thus we obtain for the unique (modulo $2\pi$) continuous function $\arg{\gamma}$ (defined by $\exp(\i\arg{\gamma})=\sgn{\gamma}$) an infinite monotone sequence $(\xi_n)_{n\in\nats}\subset I$ of distinct boundary points of these intervals such that 
\begin{equation}
|\arg{\gamma(\xi_i)}-\arg{\gamma(\xi_{i+1})}|=\pi\text{.}
\end{equation}
But this cannot happen as such sequence would converge and thus as $\arg{\gamma}$ is continuous $\lim_{i\to\infty}{|\arg{\gamma(\xi_i)}-\arg{\gamma(\xi_{i+1})}|}$.

\textbf{$2\implies1$.} Let $f$ have the desired property. Then one directly constructs a path $\gamma:I=[a,b]\to\complex\setminus\{0\}$ such that $\cfunc(\gamma)=f$. Let $\{L_i\}$ and $\{R_i\}$ be defined as in the previous part.
We then define $\gamma$ as
\begin{equation}
\gamma(x):=\exp\left(\i(\arctan{f(x)}+\pi|\{L_i\}\intersect\powerset{[a,x]}|-|\{R_i\}\intersect\powerset{[a,x]}|)\right)
\end{equation}
for $x\in I\intersect\dom(f)$. For the other points $x\in I\setminus\dom(f)$ set $\rest{\gamma}_{I\setminus\dom(f)}$ constant on path components (values $\pm\i$) such that $\gamma$ is continuous.

It is routine to check that $\cfunc(\gamma)=f$ is fulfilled. 
\end{proof}

%% remark on the connection between the pathes of a Cauchy function
\begin{remark}
Any function $\tilde{\gamma}$ with $\cfunc(\tilde{\gamma})=f$ satisfies $\tilde{\gamma}=\alpha\gamma$ where $\alpha:I\to\reels$ is continuous and of constant sign as one easily verifies. 
\end{remark}

\begin{definition}[\textsc{Cauchy} index via paths] Let $f$ be a \textsc{Cauchy} be a function on the compact interval $I=[a,b]$ corresponding to the path $\gamma:I\to\complex\setminus\{0\}$ and let $\arg{\gamma}$ be the unique (modulo $2\pi$) continuous function such that 
\begin{equation}
\gamma = \abs{\gamma}\exp(\i\arg{\gamma})\text{.}
\end{equation}
Then its \textsc{Cauchy} index is defined as
\begin{equation}
\cind(f):=\sgn(\arg{\gamma(a)}-\arg{\gamma(b)})\abs{\conv\{\arg{\gamma(a)},\arg{\gamma(b)}\}\intersect\left(\pi\ints+\frac{\pi}{2}\right)}\text{.}
\end{equation}
\end{definition}

% explanation on the definition -> rotations clockwise (traverses of imag. axis)
\begin{remark} From this we see that the Cauchy index basically counts how many times the curve $\gamma$ traverses the imaginary line clockwise, where counterclockwise traverses are counted negative (roughly speaking). 
\end{remark} % welldefined !!!

\begin{lemma}[\textsc{Cauchy} index inversion formula, most general] Let $f$ be a Cauchy function over $I=[a,b]$ and let $\{\bar{I}_i\}_{i=0}^l$ be the maximal rotational intervals of $f$. Moreover, let $\bar{I}=\conv\left(\union\{\bar{I}_i\}_{i=0}^l\right)$ be the \emph{rotational domain} of $f$ and $\bar{a}<\bar{b}$ the bounds of the interval $\bar{I}$ (which can be open or closed at each side).  
Then it holds
\begin{equation}
\cind\left(f\right)+\cind\left(\frac{1}{f}\right)=\frac{1}{2}\left(\lim_{\xi\upto\bar{b},\xi\in\bar{I}}{\sgn{f(\xi)}}-\lim_{\xi\downto\bar{b},\xi\in\bar{I}}{\sgn{f(\xi)}}\right)\text{.}
\end{equation}
\end{lemma}

\begin{proof}
Let $\tilde{\gamma}:\bar{I}\to\complex\setminus\{0\}$ be a corresponding continuous curve to $\rest{f}_{\bar{I}}$ (i.e. $\cfunc(\tilde{\gamma})=\rest{f}_{\bar{I}}$). Then $\cfunc(\i\conj{\tilde{\gamma}})=\rest{(1/f)}_{\bar{I}}$. It is clear that $\cind(\rest{f}_{\bar{I}})=\cind(f)$ from the definition of the \textsc{Cauchy} index since if $\gamma:I\to\complex\setminus\{0\}$ such that $\cfunc(\gamma)=f$ then $\gamma(a)=\tilde{\gamma}(\bar{a})$ and $\gamma(b)=\tilde{\gamma}(\bar{b})$. 
Thus in the definition of the \textsc{Cauchy} indices of $f$ and $1/f$ we get
\begin{align}
\cind(f)+\cind\left(\frac{1}{f}\right) & =\sgn(\arg{\tilde{\gamma}(\bar{a})}-\arg{\tilde{\gamma}(\bar{b})})\abs{\conv\{\arg{\tilde{\gamma}(\bar{a})},\arg{\tilde{\gamma}(\bar{b})}\}\intersect\left(\pi\ints+\frac{\pi}{2}\right)}\\
& \quad+\sgn(\arg{\i\conj{\tilde{\gamma}}(\bar{a})}-\arg{\i\conj{\tilde{\gamma}}(\bar{b})})\abs{\conv\{\arg{\delta(\bar{a})},\arg{\delta(\bar{b})}\}\intersect\left(\pi\ints+\frac{\pi}{2}\right)}\\
& = \sgn(\arg{\tilde{\gamma}(\bar{a})}-\arg{\tilde{\gamma}(\bar{b})})\abs{\conv\{\arg{\tilde{\gamma}(\bar{a})},\arg{\tilde{\gamma}(\bar{b})}\}\intersect\left(\pi\ints+\frac{\pi}{2}\right)}\\
& \quad -\sgn(\arg{\tilde{\gamma}(\bar{a})}-\arg{\tilde{\gamma}(\bar{b})})\abs{\conv\{\arg{\tilde{\gamma}(\bar{a})},\arg{\tilde{\gamma}(\bar{b})}\}\intersect\pi\ints}
\end{align}
Now set $\eta:=\max\{\arg{\tilde{\gamma}(\bar{a})},\arg{\tilde{\gamma}(\bar{b})}\}$ and $\nu:=\min\{\arg{\tilde{\gamma}(\bar{a})},\arg{\tilde{\gamma}(\bar{b})}\}$. Then we obtain from our previous calculation
\begin{align}
\cind(f)+\cind\left(\frac{1}{f}\right) & = \sgn(\arg{\tilde{\gamma}(\bar{a})}-\arg{\tilde{\gamma}(\bar{b})})\begin{cases}
1 	&: \eta\in\pi\ints+\left[-\frac{\pi}{2},0\right)\wedge\nu\in\pi\ints+\left(0,\frac{\pi}{2}\right]\\
-1 	&: \eta\in\pi\ints+\left[0,\frac{\pi}{2}\right)\wedge\nu\in\pi\ints+\left(-\frac{\pi}{2},0\right]\\
0	&: \text{ otherwise }
\end{cases}
\end{align}
Now let $L$ be the rotational interval whose lower bound is $\bar{a}$ and $U$ be the rotational interval whose upper bound is $\bar{b}$ (these exist as there are only finitely many rotational intervals as $f$ is \textsc{Cauchy} function). Then there are two cases to be distinguished at each bound namely whether $L$ or $U$ are boundary intervals (and thus half-open) or if they are open intervals. In the second case it is clear that $\arg{\tilde{\gamma}(a)}\in\pi\ints+\pi/2$ or respectively $\arg{\tilde{\gamma}(b)}\in\pi\ints+\pi/2$.

\end{proof}

\section{Some remarks on Routh-Hurwitz-Theorem}

\input{HaarSpace.tex}
\input{ChebyshevSystem.tex}
\input{FMSpace.tex}
\input{DZSpace.tex}

\section{Cauchy's Argument principle}

In this section we will deal with the question, how many zeros and poles of a meromorphic function function lie in a open set $D \subset \complex$ which is bounded by a rectifiable curve $\gamma$.

\begin{lemma}[Cauchy's Argument Principle] 
Let $f:D\to\complex$ be meromorphic inside a set $C$ and have no zeros on its contour $\gamma$. Moreover, let $P$ be the number of poles and $Z$ be the number of zeros of $f$ inside $C$. Then
%\begin{equation}
%\oint_{\gamma}{\frac{f'(z)}{f(z)}dz=2\pi\i
%\end{equation}
\end{lemma}

The theorem of Routh-Hurwitz is of remarkable importance in system and control theory taught in virtually any undergraduate course. In this section we want to take a closer look of the proof of this theorem.

\paragraph{The statement.}

To check the stability of a rational transfer function it is necessary to check weather all roots of its denominator lie in the left half-plane of $\complex$.

Such polynomials having all its roots in the left half-plane are often called \textsc{Hurwitz}-polynomials.

The check is often formulated as follows

\paragraph{Check.} Given a monic polynomial $P\in\reels[X]$ where $P=X^n+\cdots+a_0$. Such polynomial is \textsc{Hurwitz} if and only if 
\begin{enumerate}
\item All its coefficients are of the same sign (that is $\forall i\in[n] : a_i>0$).
\item All principal minors of the \textsc{Hurwitz}-Matrix $H_n(P)\in\reels^{n\timesvgfd n}$ have positive determinant where
\begin{equation}
H_n(P) := \begin{pmatrix}
a_{n-1} & a_{n-3} & \cdots & 0 & \cdots & 0\\
1 		 & a_{n-2} & \cdots & 0 & \cdots & 0\\
\vdots & \ddots & \ddots & \ddots & \ddots & \vdots\\
0 & \cdots & 0 & a_{n-1} & a_{n-3} & \cdots\\
0 & \cdots & 0 & 1 		 & a_{n-2} & \cdots\\
\end{pmatrix}\text{.}
\end{equation}
\end{enumerate}

\paragraph{The actual statement.} The theorem of \textsc{Routh-Hurwitz} as stated originally gives the above as an easy consequence. The idea is to count the difference between the number of zeros of in the left half-plane $L$ and the number of zeros in the right half-plane $R$ of $\complex$ of a polynomial $P\in\complex[X]$. 

\paragraph{The idea.}
A really simple but useful idea is here to consider the behavior of $\arg{P}$ on the imaginary axis.

To do this one needs the condition that $P$ has no root on the imaginary line. Furthermore, we assume that $P$ is monic ($a_n=1$).

Assume $P$ factors over $\complex$ such that 

\begin{equation}
P(z)=(z-l_1)\cdots(l-l_L)(z-r_1)\cdots(z-r_R)
\end{equation}

clearly this holds by fundamental theorem of algebra and by condition $R+L=n$. In the following equations let us abbreviate the operator $\lim_{\omega\to\infty}-\lim_{\omega\to-\infty}$ by $\Delta^\omega$ where $\omega\in\reels$.

\begin{equation}
\Delta^\omega{\arg{P(\omega\i)}} = \sum_{i=1}^L{\Delta^\omega{\arg(\omega\i-l_i)}}+\sum_{i=1}^R{\Delta^\omega{\arg(\omega\i-r_i)}}\label{eq1}
\end{equation}

Here $\arg$ is of course considered as continuous and takes values in $\reels$.

Let us evaluate each of the two sums separately. Thus, consider a left zero $l_i=a^l_i+b^l_i\i$ with $a^l_i<0$. One obtains that

\begin{equation}
\arg(\omega\i-l_i)=\arctan\left(\frac{\omega-b^l_i}{-a^l_i}\right)-\pi \text{ mod }2\pi
\end{equation}

As $a^l_i$ is negative, the above term increases as $\omega$ increases. Thus

\begin{equation}
\Delta^\omega{\arg(\omega\i-l_i)}=\pi
\end{equation}

and analogously for any $r_i$ we have

\begin{equation}
\Delta^\omega{\arg(\omega\i-r_i)}=-\pi\text{.}
\end{equation}

From this we get in equation (\ref{eq1}) that

\begin{equation}
\Delta^\omega{\arg{P(\omega\i)}} = \pi(L-R)\text{.}\label{eq2}
\end{equation}

\paragraph{Proceeding the argument.} A natural question is now, how we can get back from this last condition to the original polynomial. However, this is also not very difficult. Note that $P$ can be uniquely decomposed into polynomials $R$ and $S$ such that

\begin{equation}
P(\omega\i) = R(\omega)+S(\omega)\i\text{.}
\end{equation}

Now, it is a natural question to translate the term $\Delta^\omega{\arg{P(\omega\i)}}$ in a term depending on the rational function $\frac{S}{R}$ as the argument $\arg$ is connected to this function via $\arctan$.

Therefore one introduces the following definition. 

\input{CauchyIndex.tex}

\paragraph{Explanation.} The obvious intention of this definition is to count how many times a function $P(\omega\i)=R(\omega)+S(\omega)\i$ rotates around the origin $0$ of $\complex$. Therefore one may plugin $F=\frac{S}{R}$ are its reciprocal.

But we want to give some more detailed explanation.

\paragraph{Details.} %% Pictures needed! ATTENTION _ WORK TO DO
Let $\omega^{-1}_1,\ldots,\omega^{-1}_d\in\reels$ be the numbers where the function $P(\cdot\i):\reels\to\complex$ traverses the imaginary line counterclockwise and $\omega^1_1,\ldots,\omega^1_e\in\reels$ the numbers where $P(\cdot\i)$ traverses the imaginary line clockwise with respect to the origin. By condition, $P(\cdot\i)$ has no zero, thus for all these $\omega$'s we have $P(\omega\i)\neq 0$.

Formally, this means when $f:\reels\to\reels$ denotes the differentiable function $\omega\mapsto\arg{P(\omega\i)}$

\begin{equation}
R(\omega^i_j)=0 \wedge \sgn{f'(\omega^i_j)}=-i
\end{equation}

for $i=\pm 1$ and $j$ in the appropriate index set.
More, from this we see that

\begin{equation}
I_{\omega^i_j}\left(\frac{S}{R}\right)=-i\text{.}
\end{equation}

One thus may deduce the following equation

\begin{equation}
\Delta^\omega{\arg{P(\omega\i)}}=-\pi\mathcal{I}\left(\frac{S}{R}\right)+\Delta^\omega{\arctan\left(\frac{S}{R}\right)}\text{.}\label{eq4}
\end{equation}

Now, at this point, the most interesting argument comes in. Obviously, we have  $-\Delta^\omega\arg{\i\overline{P}(\omega\i)}=\Delta^\omega{\arg{P(\omega\i)}}$ and $\i\overline{P}(\omega\i)=S(\omega)+P(\omega)\i$. We thus may deduce from (\ref{eq4}) that 

\begin{align}
-\pi\mathcal{I}\left(\frac{S}{R}\right)+\Delta^\omega{\arctan\left(\frac{S}{R}\right)} = \pi\mathcal{I}\left(\frac{R}{S}\right)-\Delta^\omega{\arctan\left(\frac{R}{S}\right)}\text{.}
\end{align}

Now, let us consider the $\arctan$-terms. Indeed, here only the leading coefficients of $R$ and $S$ matter and it is easy to observe that 

BSBS
\begin{equation}
\Delta^\omega{\arctan\left(\frac{S}{R}\right)}=\pi(-1)^{\deg{R}-\deg{S}}
\end{equation}


From this definition one sees  that if $F=\frac{S}{R}$ for polynomials $R$ and $S$ then
\begin{equation}
-\pi\mathcal{I}(F)=-\pi\mathcal{I}\left(\frac{S}{R}\right)=\Delta^\omega{\arg(R(\omega)+S(\omega)\i)}\label{eq3}\text{.}
\end{equation}
 
From equations (\ref{eq2}) and (\ref{eq3}) we obtain that 

\begin{equation}
L-R = -\mathcal{I}_J\left(\frac{S}{R}\right)
\end{equation}
 


\end{document}
\message{ !name(On The Routh-Hurwitz Theorem.tex) !offset(-324) }
