\documentclass[a4paper,twoside]{article}% twoside is important for distinction of parity of sides
%%%%
\input{../Config/math.tex}
\usepackage{makeidx}
\usepackage{blindtext}
\usepackage{mathenv}
%\usepackage{showframe}% to show the frames of the page layout
\usepackage[inner=3cm,outer=2cm]{geometry} %left=4cm,right=2cm would be equivalent

% data of document
\title{Topology}
\author{Jakob Schneider}
\date{April 15, 2014}

\makeindex
\begin{document}
\pagestyle{plain}
\maketitle

\clearpage
%\pagenumbering{roman} % problem here
\pagenumbering{roman}
\tableofcontents
\pagestyle{headings}
%\clearpage
% preliminaries
\clearpage

\begin{definition}
    Let $X\in\Top$. A cover of $X$ is a set $\cU\setleq\Sub(X)$ having the following properties
    \begin{properties}
            \item $U$ covers $X$, i.e. $\Join\cU=X$.
            \item For each $x\in X$ there exists an element $U\in\cU$ such that $U\in\rest{\N}_x$, or alternatively, for all $x\in X$ we have $\rest{\N}_x X\setmeet \cU\neq\emptyset$.%
    \end{properties}
\end{definition}

\begin{definition}%%
    A space $X$ is called \keyword{compact} if for any cover $\cU$ of $X$ there is a finite subcover $\cF\in\Sub_{\fin}(\cU)$.
\end{definition}

\begin{remark}%
    An equivalent formulation is that any \emph{open} cover has the above property.
\end{remark}

\begin{definition}
    Let $X$ be a topological space. Define $\Sub_{\cpt}(X)$ as the topological space having as underlying set the compact subspaces of $X$ and as topology the sets $\gen{\set{\Sub(U)\setmeet\Sub_{\cpt}(X)}[U\in\tau]}$ where $\tau$ is the topology of $X$.
\end{definition}

\begin{remark}
    In the above system $\set{\Sub(U)\setmeet\Sub_{\cpt}(X)}[U\in\tau]$ is a basis of the topology since it is obviously closed under finite intersections.
\end{remark}

FALSE?
\begin{lemma}
    Let $\cX$ be locally compact, then $\Sub_{\cpt}{X}$ is locally compact.
\end{lemma}

\begin{proof}
    Let $C\in\Sub_{\cpt}(\cX)$ be compact. Choose compact neighboorhoods $C_c\in \rest{\N}_c\cX$ for each point of $c\in C$. Then by compactness of $C$ there is a finite subcover $\set{C_c}[c\in F]$ ($F\in\Sub_{\fin}(C)$). Define $D:=\Setjoin_{c\in F}{C_c}$ --- then $D$ is compact. We claim that $\Sub_{\cpt}(D)$ is a compact neighboorhood of $C$ in $\Sub_{\cpt}(\cX)$. It is clear that $\Sub_{\cpt}(D)\in \rest{\N}_C\Sub_{\cpt}(\cX)$ (by definition of the topology for $\Sub_{\cpt}(\cX)$). We are left to show that it is compact. To do this, choose an open cover $\cU$ of $\Sub_{\cpt}(D)$. By definition, for each $U\in\cU$ there exist open sets $V\in \tau$ such that $U=\Sub_{\cpt}(V)$. Collect these $V$ in set $\cV\setleq\tau$. As $\cU$ is an open cover of $\Sub_{\cpt}(D)$, for each compact subset $F\setleq D$ there must exist a set $V\in \cV$ such that $F\setleq V$. As the one-element subset of $D$ are compact, since $D$ is compact and $\cX$ is \person{Hausdorff} we have that $\cV$ is an open cover of $D$. ...
\end{proof}

\begin{lemma}
    Let $X$ be a locally compact space, then $X$ is a \person{Baire} space.
\end{lemma}

\begin{proof}
    Let $\set{U_n}[n\in\nats]$ be dense open subspaces of $X$ and $V$ be an open subspace of $X$. Then we define $V_0:=V$ and choose inductively $V_n$ as a non-empty relatively compact open subspace of $X$ such that $V_n\setleq \cl V_{n-1}\setmeet U_{n-1}$ for $n\geq 1$. This is possible since $U_n$ is dense in $X$ for all $n$. We thus obtain $V:=\Setmeet_{n\in\nats}{V_n}=\Setmeet_{n\in\nats}{\cl V_n}\neq\emptyset$ as a compact set such that $V\setleq \Setmeet_{n\in\nats}{U_n}\setmeet V$. This shows that $\Setmeet_{n\in\nats}{U_n}$ is dense in $X$.
\end{proof}

\begin{lemma}
    The following are equivalent
    \begin{statements}
            \item Axiom of choice.
            \item \person{Zorn}'s lemma.
        \item Wellordering theorem.
\end{statements}
\end{lemma}

\begin{proof}
    \begin{implications}
        \item[first to second] Let $P$ be preordered and for any subset $Q\setleq P$ define $Q^{\join}$ as the set of upper bounds. We then have by axiom of choice a function $F:\set{Q^{\join}}[Q\setleq P]\setminus{\set\emptyset}\to P$ with $F(X)\in X$.
    \end{implications}
\end{proof}


%\printindex%%%
%\bibliographystyle{acm}
%\bibliography{math}

\end{document}









